---
title: "MSDS_6306_CaseStudy1"
author: "Lijju Mathew"
date: "6/26/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries used

```{r}
library(plyr)
library(dplyr)
library(magrittr)
library(rlang)
library(ggplot2)
library(ggthemes)
library(knitr)
library(kableExtra)
library(class)
library(caret)
library(e1071)
library(ggthemes)
library(tidyverse)
library(GGally)
library(doBy)
library(ggmap)
library(maps)
library(openintro)
library(stringr)
library(mice)
```

# Read beer and breweries data and some initial EDA
```{r}
# Read  beer and breweries data set using read.csv
Beers <- read.csv("//Users/lijjumathew/Library/Mobile Documents/com~apple~CloudDocs/Lijju/SMU/Courses/Doing Data Science/Project/CaseStudy1_2_2_2_2_2/Beers.csv", header = TRUE, sep = ",")
Breweries <- read.csv("/Users/lijjumathew/Library/Mobile Documents/com~apple~CloudDocs/Lijju/SMU/Courses/Doing Data Science/Project/CaseStudy1_2_2_2_2_2/Breweries.csv", header = TRUE, sep = ",")
summary(Beers)
str(Beers)
summary(Breweries)
str(Breweries)
nrow(Beers)
nrow(Breweries)

detach(package:plyr)    
library(dplyr)
```

# Question 1.	How many breweries are present in each state?
```{r}
# Barchart to represent the number of breweries per state ordered
BrewPerState <-table(Breweries$State)
BrewPerState

Breweries %>% group_by(State) %>% summarize(count = n()) %>% ggplot(aes(x=reorder(State, count), y = count)) +
geom_bar(stat="identity", width=.5, fill="orange") +
labs(title="Breweries Per State Count",
     x = "State",
     y = "Brewereis Count",
     caption="source: Breweries.csv") +
     theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

# Question 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  
```{r}
# Renamed Brewery_id to Brew_ID to satisfy merging requirement
Beers <- rename(Beers, Brew_ID = Brewery_id)
BrewPub <- merge(Beers, Breweries, by="Brew_ID")
nrow(BrewPub)

# Rename Name.x as Beer_Name and Name.y as Brewery_Name
BrewPub <- rename(BrewPub, Beer_Name = Name.x)
BrewPub <- rename(BrewPub, Brewery_Name = Name.y)

# First 6 observations of the merged file
head(BrewPub, 6)
kable(head(BrewPub, 6))

# Last 6 observations of the merged file
tail(BrewPub,6)
kable(tail(BrewPub,6))
```

# Question 3.   Address the missing values in each column.
```{r}
summary(BrewPub)
MissingValues <- sapply(BrewPub, function(x)sum(is.na(x)))
# Used kable library to make document more presentable
MissingValues %>% kable("html") %>% kable_styling()

sort(sapply(who_data, function(x) sum(is.na(x))), decreasing = T)

#Missing data and percentage plot
missing.values <- BrewPub %>%
  gather(key = "key", value = "val") %>%
  dplyr::mutate(isna = is.na(val)) %>%
  dplyr::group_by(key) %>%
  dplyr::mutate(total = n()) %>%
  dplyr::group_by(key, total, isna)%>%
  dplyr::summarise(num.isna = n())%>%
  dplyr::mutate(pct = num.isna / total * 100)

levels <- (missing.values %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() + geom_bar(aes(x = reorder(key, desc(pct)), y = pct, fill=isna), stat = 'identity', alpha=1) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() + 
      labs(title = "Percentage of missing values", x = 'Variable', y = "% of missing values")
percentage.plot

# Fill missing values with mean of that column
tempData <- mice(BrewPub,m=1,maxit=0,meth='fastpmm',seed=500)
BrewPubFinal <- complete(tempData,1)

# Final check missing values
sort(sapply(BrewPubFinal, function(x) sum(is.na(x))), decreasing = T)
```



# 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
# Compute the median alcohol content unit for each state.
Alcohol <- BrewPubFinal %>% group_by(State) %>% summarize(Median = median(ABV)) %>% arrange(Median)

# Compute the median international bitterness unit for each state.
Bitter <- BrewPubFinal  %>% group_by(State) %>% summarize(Median = median(IBU)) %>% arrange(Median)

# Plot a bar chart to compare ABV by state
ggplot(data=Alcohol, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="blue")+
  ggtitle("Median ABV by State") +
  labs(x="State",y="ABV")

# Plot a bar chart to compare IBU by state
ggplot(data=Bitter, aes(x=State, y=Median)) +
  geom_bar(stat="identity", fill="blue")+
  ggtitle("Median IBU by State") +
  labs(x="State",y="IBU")
```

# Question 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
#Which state has the maximum ABV beer?  Wich state has the highest IBU beer?
BrewPubFinal[which.max(BrewPubFinal$ABV),]
kable(BrewPubFinal[which.max(BrewPubFinal$ABV),])
BrewPubFinal[which.max(BrewPubFinal$IBU),]
kable(BrewPubFinal[which.max(BrewPubFinal$IBU),])
```

# Question 6.   Comment on the summary statistics and distribution of the ABV variable.
```{r}
BeerSummary <- (summary(BrewPub$ABV))
print(BeerSummary)
ggplot(BrewPub, aes(x=ABV)) + geom_histogram(color = "blue", fill = "white") + xlab("Alcohol by Volume")
```
# When we look at summary statistics of Alcohol by volume, we see that we have some beers don't even measure 1%. Median was 5.6%, Mean or average was 5.9% and the maximum ABV was around 13%. Also, note that we are missing 62 values in the dataset that we were provided

# Looking at a histogram of the data, it seems to be a bit right-skewed.  This makes sense because beer is not meant to be a very high ABV drink.

# Question 7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}
# Draw a scatter plot to compare relationship between beer bitterness and alcohol content
ggplot(BrewPubFinal, aes(x=IBU, y= ABV)) +
  geom_point(shape=1) + 
  geom_smooth(method=lm) + # add linear regression line
  ggtitle("Correlation between IBU and ABV") +
  labs(x="IBU - Bitterness of the beer",y="ABV - Alcoholic content of the beer")

# Pearson's product-moment correlation
cor.test(BrewPubFinal$ABV,BrewPub$IBU)
```

# When we plot the ABV and IBU on a scatterplot and draw a regression line, it show evidence that there is a positive correlation.

# The Pearson correlation coefficient, r, can take a range of values from +1 to -1. A value of 0 indicates that there is no association between the two variables. A value greater than 0 indicates a positive association; that is, as the value of one variable increases, so does the value of the other variable. Here we got the cor as 0.67 between ABV and IBU which suggests there  is a positive correlation.